Disparities in Climate-Induced Health Outcomes in the Greater Toronto Area.

### 2022-06-30 outputs/paper/paper.Rmd
Updates to paper

Pollution data was collected from the Government of Canada's National Air Pollution Surveillance (NAPS) Program [website](https://data-donnees.ec.gc.ca/data/air/monitor/national-air-pollution-surveillance-naps-program/Data-Donnees/?lang=en). This website contains pollution data from 2002 to 2020. Each year contains a file with continuous data and integrated data. I will be using continuous data for each year. The continuous data contains annual summaries and hourly data. I will be looking at the hourly data for CO (Carbon Monoxide), NO (Nitric Oxide), NO2 (Nitrogen Dioxide) , O3 (Ozone), and SO2 (Sulfur Dioxide). This data includes the city (which is Toronto in our case), the levels of each pollutant for each hour of every day.

@@

-   Date

@@

temperature_2002_2022 <- read.csv("~/Desktop/SUDS2022/inputs/data/clean_temp_data/temperature_2002_2022.csv")

@@

temperature_2002_2022 %>%
  select(year, month, day, max_temp, min_temp, mean_temp) %>%
  slice(c(162, 393, 745, 1318, 1819, 2115, 3222, 4222, 5222, 6333)) %>%
  kable(
    caption = "Extracting rows from the Temperature data from 2002 - 2022",
    col.names=c("Year", "Month", "Day", "Max Temp", "Min Temp", "Mean Temp"),
    linesep = "", digits = 2, booktabs=TRUE) %>%
  kable_styling(latex_options = "HOLD_position", font_size = 10)

@@

poll_2011_2020 <- read.csv("~/Desktop/SUDS2022/inputs/data/clean_pollution_data/pollution_2011_2020.csv")

@@

poll_2011_2020 %>%
  select(Year, Pollutant, January, February, March, April, May, June, July, August, September, October, 
         November, December) %>%
  slice(c(2, 34, 74, 91, 117, 147, 184, 223, 252, 308)) %>%
  kable(
    caption = "Extracting rows from the Pollution data from 2011 - 2020",
    col.names=c("Year", "Pollutant", "Jan", "Feb", "Mar", "Apr", "May",
                "Jun", "Jul", "Aug", "Sep", "Oct", "Nov", "Dec"),
    linesep = "", digits = 2, booktabs=TRUE) %>%
  kable_styling(latex_options = "HOLD_position", font_size = 10)

@@

#### Visualizing Temperature by Months

@@

@@

>>>>

Pollution data was collected from the Government of Canada's National Air Pollution Surveillance (NAPS) Program [website](https://data-donnees.ec.gc.ca/data/air/monitor/national-air-pollution-surveillance-naps-program/Data-Donnees/?lang=en). This website contains pollution data from 2002 to 2020. Each year contains a file with continuous data and integrated data. I will be using continuous data for each year. The continuous data contains annual summaries and hourly data. I will be looking at the hourly data for CO (Carbon Monoxide), NO (Nitric Oxide), NO2 (Nitrogen Dioxide) , O3 (Ozone), and SO2 (Sulfur Dioxide). This data includes the city (which is Toronto in our case), the levels of each pollutant for each hour of every day. I reduced the data to the mean of pollution levels by pollutant of each day.

@@

-   Day

@@

temp_2002_2022 <- read.csv("~/Desktop/SUDS2022/inputs/data/clean_temp_data/temperature_2002_2022.csv")

@@

temp_2002_2022 %>%
  select(year, month, day, max_temp, min_temp, mean_temp) %>%
  slice(c(162, 393, 745, 1318, 1819, 2115, 3222, 4222, 5222, 6333)) %>%
  kable(
    caption = "Extracting rows from the Temperature data from 2002 - 2022",
    col.names=c("Year", "Month", "Day", "Max Temp", "Min Temp", "Mean Temp"),
    linesep = "", digits = 2, booktabs=TRUE) %>%
  kable_styling(latex_options = "HOLD_position", font_size = 10)

@@

pollution <- read.csv("~/Desktop/SUDS2022/inputs/data/clean_pollution_data/pollution_2002_2020.csv")

@@

pollution %>%
  select(Pollutant, City, Date, Mean, Year) %>%
  slice(c(100, 1000, 1200, 1500, 1800, 2000, 2500, 2800, 3600, 5000, 6000, 7000, 10000, 11000, 15000)) %>%
  kable(
    caption = "Extracting rows from the Pollution data from 2002 - 2020",
    col.names=c("Pollutant", "City", "Date", "Mean", "Year"),
    linesep = "", digits = 2, booktabs=TRUE) %>%
  kable_styling(latex_options = "HOLD_position", font_size = 10)

@@

#### Visualizing Temperature by Month and Year

@@

In the above plot...

@@

In the above plot...

## Visualizing Pollution from 2002-2020

#### Visualizing Pollution by Year

```{r echo=FALSE, message=FALSE, warning=FALSE, fig.cap="Mean Pollution 2002 to 2020 for NO, NO2, O3, SO2 by Year", fig.width=9, fig.height=3.2}

pollution %>% filter(pollutant != "CO") %>% ggplot(
  aes(x = year, y = mean_pollution, color = pollutant)) + geom_point() + 
  labs(color="Pollutant") + xlab("Year") + ylab("Mean Pollution (ppb)")
```

In the above plot...

```{r echo=FALSE, message=FALSE, warning=FALSE, fig.cap="Mean Pollution 2002 to 2020 for CO by Year", fig.width=9, fig.height=3.2}

pollution %>% filter(pollutant == "CO") %>% ggplot(
  aes(x = year, y = mean_pollution, color = pollutant)) + geom_point() + 
  labs(color="Pollutant") + xlab("Year") + ylab("Mean Pollution (ppm)")

In the above plot...

#### Visualizing Pollution by Month

```{r echo=FALSE, message=FALSE, warning=FALSE, fig.cap="Mean Pollution 2002 to 2020 for NO, NO2, O3, SO2 by Month", fig.width=9, fig.height=3.2}

pollution %>% filter(pollutant != "CO") %>% ggplot(
  aes(x = month_name, y = mean_pollution, color = pollutant)) + geom_point() + 
  labs(color="Pollutant") + xlab("Month") + ylab("Mean Pollution (ppb)")

In the above plot...

```{r echo=FALSE, message=FALSE, warning=FALSE, fig.cap="Mean Pollution 2002 to 2020 for CO by Month", fig.width=9, fig.height=3.2}

pollution %>% filter(pollutant == "CO") %>% ggplot(
  aes(x = month_name, y = mean_pollution, color = pollutant)) + geom_point() + 
  labs(color="Pollutant") + xlab("Month") + ylab("Mean Pollution (ppm)")

In the above plot...

## Visualizing Pollution and Temperature from 2002-2020

#### Visualizing Mean Pollution by Temperature, Pollutant and Month

```{r echo=FALSE, message=FALSE, warning=FALSE, fig.cap="Mean Pollution 2002 to 2020 for NO, NO2, O3 by Month, Maximum Temperature, and Pollutant for December to March", fig.width=9, fig.height=3.2}

pollu_temp %>% filter(pollutant != "CO" & pollutant != "SO2" & 
  (month_name == "Dec" | month_name == "Jan"| month_name == "Feb" | 
  month_name == "Mar")) %>% ggplot(aes(max_temp, mean_pollution, 
  color=month_name)) + geom_point() + facet_grid(. ~ pollutant) + 
  labs(color="Month") + xlab("Maximum Temperature") + 
  ylab("Mean Pollution")
```

In the above plot...

```{r echo=FALSE, message=FALSE, warning=FALSE, fig.cap="Mean Pollution 2002 to 2020 for CO and SO2 by Month, Maximum Temperature, and Pollutant for December to March", fig.width=9, fig.height=3.2}

pollu_temp %>% filter((pollutant == "CO" | pollutant == "SO2") & 
  (month_name == "Dec" | month_name == "Jan"| month_name == "Feb" | 
  month_name == "Mar")) %>% ggplot(aes(max_temp, mean_pollution, 
  color=month_name)) + geom_point() + facet_grid(. ~ pollutant) + 
  labs(color="Month") + xlab("Maximum Temperature") + 
  ylab("Mean Pollution")

In the above plot...

```{r echo=FALSE, message=FALSE, warning=FALSE, fig.cap="Mean Pollution 2002 to 2020 for NO, NO2, O3 by Month, Maximum Temperature, and Pollutant for April to July", fig.width=9, fig.height=3.2}

pollu_temp %>% filter(pollutant != "CO" & pollutant != "SO2" & 
  (month_name == "Apr" | month_name == "May" | month_name == "Jun" | 
  month_name == "Jul")) %>% ggplot(aes(max_temp, mean_pollution, 
  color=month_name)) + geom_point() + facet_grid(. ~ pollutant) + 
  labs(color="Month") + xlab("Maximum Temperature") + 
  ylab("Mean Pollution")

In the above plot...

```{r echo=FALSE, message=FALSE, warning=FALSE, fig.cap="Mean Pollution 2002 to 2020 for CO and SO2 by Month, Maximum Temperature, and Pollutant for April to July", fig.width=9, fig.height=3.2}

pollu_temp %>% filter((pollutant == "CO" | pollutant == "SO2") & (month_name == "Apr" |
   month_name == "May" | month_name == "Jun" | month_name == "Jul")) %>% ggplot(aes(max_temp,
   mean_pollution, color=month_name)) + geom_point() + facet_grid(. ~ pollutant) + 
  labs(color="Month") + xlab("Maximum Temperature") + 
  ylab("Mean Pollution")

In the above plot...

```{r echo=FALSE, message=FALSE, warning=FALSE, fig.cap="Mean Pollution 2002 to 2020 for NO, NO2, O3 by Month, Maximum Temperature, and Pollutant for August to November", fig.width=9, fig.height=3.2}

pollu_temp %>% filter(pollutant != "CO" & pollutant != "SO2" & (month_name == "Aug" | 
    month_name == "Sep" | month_name == "Oct" | month_name == "Nov")) %>% ggplot(aes(max_temp,
    mean_pollution, color=month_name)) + geom_point() + facet_grid(. ~ pollutant) + 
    labs(color="Month") + xlab("Maximum Temperature") + ylab("Mean Pollution")

In the above plot...

```{r echo=FALSE, message=FALSE, warning=FALSE, fig.cap="Mean Pollution 2002 to 2020 for CO and SO2 by Month, Maximum Temperature, and Pollutant for August to November", fig.width=9, fig.height=3.2}

pollu_temp %>% filter((pollutant == "CO" | pollutant == "SO2") & 
  (month_name == "Aug" | month_name == "Sep" | month_name == "Oct" | month_name == "Nov")) %>%
  ggplot(aes(max_temp, mean_pollution, color=month_name)) + geom_point() + facet_grid(. ~
  pollutant) + labs(color="Month") + xlab("Maximum Temperature") + ylab("Mean Pollution")

In the above plot...
